Learning and Evaluating \strips\ Action Models with Classical Planning

### 2018-03-20 doc/ijcai-18/rebuttal/rebuttal.txt

Rebuttal #1:
-------------
Unlike extensive-data ML, this work explores an alternative research direction to learn sound models from small amounts of correct input data. Further, this is the first work on learning and evaluating STRIPS action models from state observations that report reproducible results over a wide range of domains exclusively using an off-the-shelf classical planner. We belief this approach is relevant by itself because opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.

Rebuttal #2:
-------------
We agree with the reviewer that a SAT-based compilation could be used for the addressed learning task. With this regard, the SAT-based system ARMS is the closest system to compare with (also requires the operator heads). ARMS  addresses however a different learning task: (1) learns from plan traces (the actions observations are assumed to be correct) while we learn from state observations and (2), ARMS addresses an optimization task (i.e. maximizing the number of covered examples) while ours is a satisfying task, we compute a conservative model that is guaranteed to be applicable at ALL the given observations.

Like happens with satisfying planners, our approach is extensible to address the corresponding optimization task, e.g. in our case minimizing the number of model edits required to cover a set of partially-observable/noisy state observations. This work aimed at understanding the grounds of the compilation for learning and evaluating STRIPS models in fully-observable environments (e.g. classical planning tasks, video-games, simulators) but extending this approach to partially-observable/noisy environments is in our research agenda. With this said, we bet on classical planning because we believe that existing relaxations of the classical planning task, such as the "Relaxed plan" have an important role to play in the assessment of generative models with regard to large amounts of partially-observable/noisy examples. Further, when there is an unknown number of missing observations, planning seems more natural than SAT because the planning horizon is no longer known.

Answering the reviewer questions:
1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) because we obtained better experimental results (we attach the results we got when starting from the most general hypothesis (no preconditions). 
2. We agree with the reviewer that the "one test_i fluent" is a more natural encoding. We will evaluate it to see if we achieve some performance gain.

Rebuttal #3:
-------------
The contribution of our approach is leveraging classical planning not only for (1) LEARNING, that allows us to report quantitative results over a wide range of different domains but also for (2), EVALUATING the quality of a STRIPS model with regard to a test set of observations.

Evaluating STRIPS models with a test set of observations enables the model recognition task. This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape). Different generative models like policies, programs or grammars are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al). Therefore, this work poses a general method to assess how a given generative model matches a test set of observations: for a wide range of different generative models (the STRIPS compilable ones), and exclusively using classical planning. As far as we know, this is the first work on doing so.

Rebuttal #4:
-------------
The empirical evaluation focused on understanding which kind of STRIPS action models could be learned from just 25 state observations. The results revealed that 25 state observations were enough to learn action models with few action schemes,  (e.g "blocks" or "gripper"). In some of these domains ("hanoi", "visitall", "npuzzle") the reported Precision&Recall did not reach 1.0 for the preconditions. The reason is that learning discovered preconditions that were not specified in the actual/reference domain. For instance, our approach learned for the "move" action in the "visitall" the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1), while only the (connected ?place1 ?place2) precondition is present in the actual/reference model, despite both preconditions are actually correct.

Learning accurate models from few examples becomes harder in domains with larger number of action schemes, in these cases the different actions have lower chances of being chosen. Examples are the "disembark-truck" action missing from the "driverlog" domain, the "paint-down" action missing from the "floortile" domain or the "move-curb-to-curb" from the "parking" domain. Including more examples, specially examples of these missing actions, definitely leads to better strips action models.

Because our approach is a classical planning compilation, the limit on the number of input observations is given by the limitation of the planner performance. In the particular case of current SAT-based planners, they are known to have performance issues with planning horizons beyond the 150-200 steps (this would currently limit our approach to about a hundred of input examples).

With this said, we agree with the reviewer that the set of input problems definitely affects to the quality of the learned models. Currently were are selecting problems for each domain with a small number of objects because they are easier to be solved and the result of the compilation is more tractable. An interesting research question is identifying which is the smallest number of objects per type required to learn a complete STRIPS action model for a given domain. This question has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, J. Rintanen IJCAI 2017). We believe that this research direction is promising to obtain similar results for learning STRIPS action models.

### 2018-03-21 doc/ijcai-18/rebuttal/rebuttal.txt

Rebuttal #1:
-------------
We sincerely thank reviewers for their constructive comments and criticisms which help to clarify details of our approach. Unlike extensive-data ML, our approach explores an alternative research direction to learn sound action models from a small amount of correct input data. Further, this is the first work that learns and evaluates action models exclusively using an off-the-shelf classical planner. We belief this result is relevant by itself because it opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.

The SAT-based system ARMS is the closest system to compare with (also requires the operator heads). ARMS addresses however a different learning task: (1) learns from plan traces (the actions observations are assumed to be correct) while we learn from state observations (2), ARMS addresses an optimization task (i.e. maximizing the # of covered examples) while ours is a satisfying task, we compute a conservative model that is guaranteed to be applicable at ALL the given observations and (3), ARMS evaluates the quality of the learned models using two different functions "error rate" that evaluates the quality of the learned preconditions and "redundancy" that evaluates the quality of the positive effects.

Rebuttal #2:
-------------
In effect a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper.  We bet on classical planning because when there is an unknown # of missing observations, planning seems more natural than SAT because the planning horizon is no longer known and addressing these tasks is in our research agenda. Further, the planning commmunity has shown a tremendous experince in the develpment of informative while inexpensiver heuristic functions able to evaluate millions of symbolic states in seconds time. We believe that existing relaxations of the classical planning task, such as the "Relaxed plan" have an important role to play in the assessment of generative models with regard to large amounts of partially-observable/noisy examples.

Like happens with satisfying planners, our approach is extensible to address the corresponding optimization task, e.g. in our case minimizing the # of model edits required to cover a set of partially-observable/noisy state observations. As a first step, this work aimed at understanding the grounds of the compilation for learning and evaluating STRIPS models in fully-observable environments (e.g. classical planning tasks, video-games, simulators) but extending this approach to partially-observable/noisy environments is definitely in agenda.

Answering the reviewer questions:
1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained better experimental results. 
2. We agree with the reviewer that the "one test_i fluent" is a more natural encoding. We plan to evaluate and chekc whether we achieve some performance gain.

Rebuttal #3:
-------------
The contribution of our approach is leveraging classical planning not only for (1) LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for (2), EVALUATING how well a given STRIPS model matches a test set of observations.

Evaluating STRIPS models according to a test set of observations enables the "STRIPS model recognition" task. This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape). Different generative models like policies, programs, grammars and different forms of domain-specific control knowledge are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al). With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a test set of observations.

Rebuttal #4:
-------------
The evaluation focused on understanding which models can be learned from just 25 observations. Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas (this # ranges [1,7] in our evaluation). The largest this # the lowest chance an operator of being chosen (e.g. "disembark-truck" action is missing from the "driverlog" domain or "paint-down" is missing from "floortile"). Including more observations, specially of the missing actions, leads to better action models. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SAT-based planners have performance issues with planning horizons beyond 150-200 steps (since actions for programming preconditions/effects are often applied in parallel, in a single step, this means we can handle 50-80 input observations with a SAT-based planner.

Precision&Recall did not reach 1.0 in some domains with few schemes because learning "discovered" preconditions that were not specified in the IPC/reference domain. For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in "visitall", while only (connected ?place1 ?place2) was specified in the IPC/reference model.

The input observations affects to the quality of the learned models. Were are using one problem for each domain with a small # of objects because the compiled task is more tractable. If there is IPC instances small enough, we use an IPC instance, otherwise we created a small problem. An interesting research question is identifying which is the smallest # of objects per type required to learn a complete STRIPS action model for a given domain. This question has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, J. Rintanen IJCAI 2017). We believe that this research direction is promising to obtain similar results for learning STRIPS action models.

### 2018-03-22 doc/ijcai-18/rebuttal/rebuttal.txt

Rebuttal #12637:
-------------
We sincerely thank reviewers for their constructive comments and criticisms which help to clarify details of our approach. This is the first work that LEARNS and EVALUATES action models exclusively using an off-the-shelf classical planner. We belief this result is relevant by itself because it opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.

The SAT-based system ARMS (AIJ07) is the closest system to compare with, since ARMS also requires the operator heads and observations of correct plans. ARMS addresses however a different learning task:
(1) ARMS learns from plan traces (where the ACTIONS observations are assumed to be correct) while we learn from STATE observations.
(2) ARMS addresses an OPTIMIZATION task (i.e. maximizing the # of covered examples) while ours is a SATISFYING task. We compute a conservative model that is guaranteed to be applicable at all the given observations.

Rebuttal #23413:
-------------
In effect, we definitely agree with the reviewer that a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper. The only reason why we bet on planning is because, when there is an unknown # of missing observations, the planning horizon is no longer known and addressing this learning task is in our research agenda. Further, the planning community has shown terrific expertise in the development of informative while inexpensive heuristics (current heuristics can evaluate millions of symbolic states in seconds time). Our believe is that this technology has a key role to play in the assessment of generative models with regard to large amounts of incomplete/noisy examples.

Like happens with satisfying planners, our approach is extensible to the corresponding optimization task, e.g. in our case minimizing the # of model edits required to cover a set of observations. As a first step, this paper aimed at understanding the grounds of our compilation for learning and evaluating STRIPS models in fully-observable environments (e.g. classical planning tasks) but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.

Answering the reviewer questions:
1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results. There is nothing that prevent us from starting with the most general hypothesis (zero preconditions). In fact the evaluation task starts with a given action model that may have for instance, no preconditions. 
2. We agree with the reviewer that the "one test_i fluent" encoding is more natural. We plan to evaluate it and check whether it allow us to achieve some performance gain.
To comments on the training set we refer the reviewer to the Rebuttal #26170.

Rebuttal #25749:
-------------
The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for EVALUATING how well a given STRIPS model matches a test set of observations.

Evaluating a given STRIPS model according to a test set of observations enables the "STRIPS model recognition" task. This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape). Different generative models like policies, programs, grammars and different forms of domain-specific control knowledge are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al). With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a given set of observations.

Rebuttal #26170:
-------------
Our evaluation focused on understanding which models can be learned from just 25 observations. Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas. In our evaluation this # ranges [1,7]. The largest this # the lowest chance of choosing an operator, e.g. "disembark-truck" is missing from the driverlog or "paint-down" is missing from floortile. Including more observations, specially of the missing schemes, leads to better models. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SATplanners have performance issues with horizons beyond 150-200 steps. Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (we require 2 actions to validate 1 observation, Fig 3).

Precision&Recall did not reach 1.0 in some domains with few schemes because learning discovered preconditions that were not specified in the IPC/reference domain. For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, while only (connected ?place1 ?place2) was specified in the IPC model. Effects were perfectly learned in these domains.

The input observations affect to the quality of the learned models. Were use 1 problem for each domain with a small # of objects because this way the compiled task is more tractable. If IPC instances are small enough we use them otherwise, we created a smaller problem. An open question is identifying the smallest # of objects per type required to learn a complete STRIPS model. This has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, Rintanen IJCAI17). We believe this research direction is promising to obtain similar results for learning STRIPS models.

### 2018-03-22 doc/ijcai-18/rebuttal/rebuttal_March22_Sergio_Diego_Eva.txt

Rebuttal #12637:
-------------

We would like to thank all the reviewers for their thoughtful comments and insights as well as expressing our general agreement with the provided comments.

Lack of comparison
----------------------

The SAT-based system ARMS (AIJ07) is the closest system to compare with since ARMS also requires the operator heads and observations of correct plans. ARMS addresses, however, a different learning task:

(1) ARMS learns from plan traces (where the observed ACTIONS are assumed to be correct) while we learn from STATE observations (unobserved actions).
(2) ARMS addresses an OPTIMIZATION task (i.e. maximizing the # of covered examples) while ours is a SATISFYING task. We compute a conservative model that is guaranteed to be applicable at all the given observations.

As for LOCM, we think a better comparison would be with LOP since LOCM only learns the dynamic aspects of the domain (i.e. state changes that occur due to action application but not the static relations). However, LOP needs a set of optimal plans as input to the learning system.

Rebuttal #23413:
-------------
We definitely agree with the reviewer that a SAT-based compilation could also be used for learning and evaluating the tasks addressed in this paper. The only reason why we opt for planning is because when there is an unknown # of actions between state observations, the planning horizon is no longer known and this learning task is in our research agenda. Further, the planning community has shown terrific expertise in the development of informative and inexpensive heuristics (current heuristics are able to evaluate millions of symbolic states in seconds). Our belief is that this technology will play a key role in the assessment of generative models with inputs of large amounts of incomplete/noisy examples.

Like happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations). As a first step, this paper aims at understanding the grounds of the compilation scheme for learning and evaluating models in fully-observable environments (e.g. classical planning tasks); but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.

Answers to the reviewer's questions:

1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results. Nothing prevents us from starting with the most general hypothesis (zero preconditions), a random model, or even a previously learnt model (as it is the case of the evaluation task). In fact, the evaluation task starts with a given action model that may contain, for instance, no preconditions at all.

2. We thank the reviewer for the suggestion that the "one test_i fluent" encoding is more natural. We will evaluate it and check if this brings some performance gain.

Regarding the remarks on how the training set was generated, we refer the reviewer to the Rebuttal #26170.

Rebuttal #25749:
-------------
The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains but also for EVALUATING how well a given STRIPS model matches a test set of observations.

Evaluating a given model according to a test set of observations represents a step towards solving the "model recognition" task. Given the high expressivenes of the STRIPS models (as expressive as a Turing Machine with finite tape), the task of model recognition is certainly relevant. Different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable (as stated in the series of work by Jorge Baier, Sheila McIlraith et al). In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) in a given set of observations.

Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to the Rebuttal #12637.

Rebuttal #26170:
---------------

Generation of the training set
------------------------------

The 25 observations per domain were generated by solving one instance of the IPC; in some cases we modified the problem and limited the number of objects to keep the computation time low. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SATplanners have performance issues with horizons beyond 150-200 steps. Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (note that we require 2 actions to validate one observation, Fig 3). Ultimately, our goal was to evaluate how well the approach performs with few observations and so it is more important to have representative samples than a large number of observations, which would undoubtedly lead to better models but also longer learning times.

Given the small size of the training set, there is no guarantee that the unobserved plan contains actions for all the domain schemas, which clearly has an impact in some results (e.g, the missing actions “disembark-truck” and “paint-down” of the driverlog and floortile domains, respectively). Another reason that explains that domains are not perfectly learnt (wrt the reference domain) is the appearance in the output models of preconditions that although correct they are usually skipped in the domain definition because of redundancy. For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, but only (connected ?place1 ?place2) is specified in the IPC model.

An open question is the identification of the smallest # of objects per type required to learn a complete STRIPS model. This has recently been addressed for Schematic Invariants (Rintanen IJCAI17). We believe this is a promising research direction to obtain similar results in learning STRIPS models.

Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to the Rebuttal #12637.

### 2018-03-23 doc/ijcai-18/rebuttal/rebuttal.txt

Rebuttal #12637:
-------------

We would like to thank all the reviewers for their thoughtful comments and insights as well as expressing our general agreement with the provided comments.

The SAT-based system ARMS (AIJ07) is the closest system to compare with since ARMS also requires the operator heads and observations of correct plans. ARMS addresses, however, a different learning task:

(1) ARMS learns from plan traces (where the observed ACTIONS are assumed to be correct) while we learn from STATE observations (the executed actions are unobserved).
(2) ARMS addresses an OPTIMIZATION task (i.e. maximizing the # of covered examples) while ours is a SATISFYING task. We compute a conservative model that is guaranteed to be applicable at all the given observations.

As for LOCM, we think a better comparison would be with LOP since LOCM only learns the dynamic aspects of the domain (i.e. state changes that occur due to action application but not the static relations). However, LOP needs a set of optimal plans as input to the learning system.

Rebuttal #23413:
-------------
We definitely agree with the reviewer that a SAT compilation could also be used for learning and evaluating the tasks addressed in this paper. The only reason why we opt for planning is because when there is an unknown # of actions between state observations, the planning horizon is no longer known and this learning task is in our research agenda. Further, the planning community has shown terrific expertise in the development of informative/inexpensive heuristics (current heuristics can evaluate millions of symbolic states in seconds). Our belief is that this technology can play a key role in the assessment of generative models with large amounts of incomplete/noisy examples.

Like happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations). As a first step, this paper aims at understanding the grounds of the compilation scheme for learning and evaluating models in fully-observable environments (e.g. classical planning tasks); but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.

Answers to the reviewer's questions:
1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results. Nothing prevents us from starting with the most general hypothesis (zero preconditions), a random model, or even a previously learnt model (as it is the case of the evaluation task). In fact, the evaluation task starts with a given action model that may contain, for instance, no preconditions at all.

2. We thank the reviewer for the suggestion. We agree that the "one test_i fluent" encoding is more natural. We will evaluate it and check if this brings some performance gain.

Regarding the remarks on how the training set was generated, we refer the reviewer to our rebuttal to Review  #26170.

Rebuttal #25749:
-------------
The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains but also for EVALUATING how well a given STRIPS model matches a test set of observations.

Evaluating a given model according to a test set of observations represents a big step towards the "model recognition" task. Given the high expressiveness of the STRIPS models (as expressive as a Turing Machine with finite tape), the task of model recognition is certainly relevant. Different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable (as stated in the series of work by Jorge Baier, Sheila McIlraith et al). In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.

Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to our rebuttal to Review #12637.

Rebuttal #26170:
---------------

We generated 25 training observations per domain by solving 1 IPC instance per domain; in some cases we modified that instance, limiting the number of objects, to keep computation time low. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SATplanners have performance issues with horizons beyond 150-200 steps. Given that actions for programming preconds/effects are often applied in parallel (in 1 step) this means we could handle no more than a hundred observations with current SATplanners (recall we require 2 actions to validate 1 observation, Fig 3).

Given the small size of the training set, there is no guarantee that the unobserved plan contains actions for all the domain schemas, which clearly affects some results (e.g, actions 'disembark-truck' and 'paint-down' are missing from the driverlog and floortile domains, respectively). We aimed evaluating how well our approach performs with just 25 observations. A larger # of observations, specially of the missing actions, would lead to better models but also longer learning times.

Another reason that explains that models are not perfectly learnt (wrt the IPC domains) is the appearance in the learnt models of preconds that although correct, are skipped in the IPC models because of redundancy. For instance, we learned preconds (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, but only (connected ?place1 ?place2) is specified in the IPC model.

An open question is identifying the smallest # of objects per type required to learn a complete STRIPS model. This has recently been addressed for Schematic Invariants (Rintanen IJCAI17). We believe this is a promising research direction to obtain similar results for learning STRIPS models. Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to our rebuttal to Review #12637.

### 2018-03-23 doc/ijcai-18/rebuttal/rebuttal_March22_Sergio_Diego_Eva.txt

(1) ARMS learns from plan traces (where the observed ACTIONS are assumed to be correct) while we learn from STATE observations (unobserved actions).
(2) ARMS addresses an OPTIMIZATION task (i.e. maximizing the # of covered examples) while ours is a SATISFYING task. We compute a conservative model that is guaranteed to be applicable at all the given observations.

As for LOCM, we think a better comparison would be with LOP since LOCM only learns the dynamic aspects of the domain (i.e. state changes that occur due to action application but not the static relations). However, LOP needs a set of optimal plans as input to the learning system.

Rebuttal #23413:
-------------
We definitely agree with the reviewer that a SAT-based compilation could also be used for learning and evaluating the tasks addressed in this paper. The only reason why we opt for planning is because when there is an unknown # of actions between state observations, the planning horizon is no longer known and this learning task is in our research agenda. Further, the planning community has shown terrific expertise in the development of informative and inexpensive heuristics (current heuristics are able to evaluate millions of symbolic states in seconds). Our belief is that this technology will play a key role in the assessment of generative models with inputs of large amounts of incomplete/noisy examples.

As it happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations). As a first step, this paper aims at understanding the grounds of the compilation scheme for learning and evaluating models in fully-observable environments (e.g. classical planning tasks); but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.

@@

Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to the Rebuttal #12637.

Rebuttal #26170:
---------------

Generation of the training set
------------------------------

The 25 observations per domain were generated by solving one instance of the IPC; in some cases we modified the problem and limited the number of objects to keep the computation time low. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SATplanners have performance issues with horizons beyond 150-200 steps. Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (note that we require 2 actions to validate one observation, Fig 3). Ultimately, our interest was to evaluate the scope of the compilation scheme when learning with few observations. Thereby, having representative samples is more relevant to us than having a large number of observations, which would undoubtedly lead to better models but also longer learning times.

### 2018-03-23 doc/aij18/aij18.tex
policy/goal learning added to aij18 paper

\subsection{Learning from observations of plan executions}
The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its plan executions, the actual executed actions are unobserved. This learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
\begin{itemize}
\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$. In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes. This means that some preconditions and some positive or negative effects are a priori known.
\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The states in $\mathcal{O}$ are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (it is unknown whether their value is either positive or negative). Anyway we assume that observations are noiseless, meaning that if the value of a fluent is observed it is correct.
\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. This set of predicates can be given as input but it can also be inferred from the state observations given in $\mathcal{O}$ provided that they are full states.
\end{itemize}

@@

>>>>

\subsection{Learning from observations of plan executions}
The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its plan executions, the actual executed actions are unobserved. This learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
\begin{itemize}
\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$. 
\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The states in $\mathcal{O}$ are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (it is unknown whether their value is either positive or negative). Anyway we assume that observations are noiseless, meaning that if the value of a fluent is observed it is correct.
\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. This set of predicates can be given as input but it can also be inferred from the state observations given in $\mathcal{O}$ provided that they are full states.
\end{itemize}

@@

\subsection{Learning from a partially specified model}
In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes. This means that some preconditions and some positive or negative effects are a priori known.

Such scenario is specially relevant to addressing two well-known learning tasks:
\begin{itemize}
\item {\em Policy learning}. A policy is function that maps states into actions and represent the conditions under which an action should be applied to achieve certain goals. Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each of the action schemes $\xi\in\mathcal{M}$. These extra preconditions capture when actions can be applied according to the policy. If for each possible state there is only one applicable action we say that the policy is fully specified otherwise, we say that the policy is partially specified. In the general case, policy learning for arbitrary planning tasks is a complex task because it may require computing new {\em high-level state features}~\cite{lotinac2016automatic}.

\item {\em Goal learning}. The goal condition is the set of constraints that defines the subset of states that are considered goal states. Given an action model $\mathcal{M}$ the task of learning the goal conditions for a a given set of observations of plan executions can be defined as learning the preconditions of a extra action that is always applied in the last place and whose only effect is marking that the goals of the planning task are reached.
\end{itemize}
